---
title: "2021 144L Abundance Laptopv"
author: "Ben Caulfield"
date: "10/18/2021"
output: github_document
---

**Currently unable to knit due to an issue opening image files that the script itself makes.

# Goal
Analyze Bacterial abundance data from 2021 remineralization experiments.

```{r Load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

```{r importing}
excel_sheets("~/R/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx")


metadata <- read_excel("~/R/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx")

glimpse(metadata)
#unique(metadata$Bottle)
#unique(metadata$Treatment)

data <- read_excel("~/R/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx", sheet = "Data")
glimpse(data)
joined <-  left_join(metadata, data) #attach data to metadata
glimpse(joined)

```

```{r Cells data prep, warning=FALSE, message=FALSE}
cells <- joined %>%
  mutate(Datetime = ymd_hm(Datetime), #splits apart Datetime as specified
  
  cells_L =  as.numeric(all_cells_uL)* 1000000) %>%
  group_by(Treatment, Bottle) %>%
#group our dataset so that we can calculate the time elapsed properly 
  mutate(interv = interval(first(Datetime), Datetime), 
         s = as.numeric(interv), 
         hours = s/3600, 
         days = hours/24) %>%
  ungroup() %>%
  select(Experiment:DNA_Sample, cells_L, hours, days) %>%
  drop_na(cells_L)
glimpse(cells)
```

```{r Initial Plot Growth Curves,warning=FALSE}

custom.colors <- c("Control" = "#66c2a5", "Kelp Exudate_Nitrate_Phosphate" = "#fc8d62", "Kelp Exudate" = "#8da0cb", "Glucose_Nitrate_Phosphate" = "#e78ac3")
levels <- c("Control", "Kelp Exudate_Nitrate_Phosphate", "Kelp Exudate", "Glucose_Nitrate_Phosphate")
cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=cells_L, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size =1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) + 
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  #facet_grid(rows = "Treatment")
  theme_bw()
glimpse(cells)
```

```{r ln & diff_ln creation}
ln_cells <- cells %>%
  group_by(Treatment, Bottle) %>%
  mutate(ln_cells = log(cells_L), 
         diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells)))
```

```{r Plotting ∆ln_cells, fig.height=6, fig.width=8, warning=FALSE}
ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=diff_ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size =1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) + 
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Difference ln cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle", ncol =2)+
  #facet_grid(rows= "Treatment")
  
  theme_bw()
```


```{r Plotting ln_cells, fig.height=6, fig.width=8, warning=FALSE}
ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size =1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) + 
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("ln cells, L"^-1)), fill = "") + 
  guides(color = "none") + 
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle", ncol =2) +
  theme_bw()
```

*NOTE

Some of the abundance timepoints are missing. this is due to sample labels falling off in LN2. This should be fixed by 10/25/21 since the groups with the wonky data went and remade the samples? Unless I'm dumb and they were just doing the PCR again.

Three types of abundance data: all_cells, LNA_cells (low nucleic acid) and HNA_cells (high nucleic acid). you may want to investigate trends present in all three.

